---
title: "Coho CWT FRAM & MSM Analysis"
author: "Dan.Auerbach@dfw.wa.gov, Angelika.Hagen-Breaux@dfw.wa.gov, carrie_cook-tabor@fws.gov"
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
output: 
  bookdown::html_document2:
    fig_caption: yes
    theme: cerulean
    toc: yes
    toc_depth: 3
    toc_float: yes
    code_folding: hide
---

```{r setup, results = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, results = FALSE, warning = FALSE, message = FALSE, fig.width = 9, fig.height = 9)

library("tidyverse")
library("gt")
library("odbc"); library("DBI")
#library("ggridges")

fp <- list(
  db_msm = "T:/DFW-Salmon Mgmt Modeling Team - General/Southern Fund Projects/Coho FRAM MSM CWT analysis/MSMDB/MSM_98-2020- SF2019.mdb"
  # data_fram = "data/postseason_fram_spwn_hvst_ests.csv",
  # data_cwt = "data/rmis_cwts.csv",
  )

```

# Summary

# Data

```{r read_in_rmis_recoveries, eval=FALSE}
dir_rmis_recs <- "T:/DFW-Salmon Mgmt Modeling Team - General/Southern Fund Projects/Coho FRAM MSM CWT analysis/ERAnalysis/RMISRecoveries/"

focal_cols_recs <- c("recovery_id", "tag_code", "run_year", "recovery_location_code", "recovery_location_name", "fishery", "gear", "estimated_number")

# rmis_recs <- purrr::map_dfr(
#   list.files(dir_rmis_recs, full.names = T)
#   ,
#   ~readr::read_csv(.x, col_select = all_of(focal_cols_recs))
#   )
# 
# saveRDS(rmis_recs, "rmis_coho_recoveries_86_2020.rds")

```

```{r}
rmis_recs <- readRDS("rmis_coho_recoveries_86_2020.rds")

rmis_recs |> 
  count(estimated_number < 1) |> 
  mutate(pct_n = n / nrow(rmis_recs))

rmis_recs |> count(fishery) |> arrange(desc(n)) |> print(n=100)

#everything
rmis_recs |> count(run_year) |> ggplot(aes(run_year, n)) + geom_col() + geom_smooth()
#exclude NA estimated number...not a huge diff
rmis_recs |> 
  filter(!is.na(estimated_number)) |> 
  count(run_year) |> 
  ggplot(aes(run_year, n)) + geom_col() + geom_smooth()

rmis_recs |> 
  mutate(across(c(fishery, gear), factor)) |> 
  count(run_year, fishery) |> 
  ggplot(aes(run_year, fishery)) +
  geom_raster(aes(fill = n))

rmis_recs |> 
  mutate(
    f_type = case_when(
      between(fishery, 10, 19) ~ "troll",
      between(fishery, 20, 29) ~ "net",
      between(fishery, 40, 49) ~ "sport",
      between(fishery, 50, 59) ~ "escp"
    )
    ) |> 
  count(run_year, f_type) |> 
  #ggplot(aes(run_year, f_type)) +
  # geom_raster(aes(fill = n)) +
  # scale_fill_distiller(direction = 1)
  ggplot(aes(run_year, n, fill = f_type)) +
  geom_col() +
  facet_wrap(~f_type)

```


[NEEDS REVISION]
After querying RMIS for coho recoveries by year, the raw recovery files are loaded into an Access database, which is then used to run the Mixed Stock Model (MSM). MSM associates individual recoveries to fisheries via the recovery code and gear fields and various lookups. The output of that mapping is exported as delimited files by year (a)


```{r}
#86-91, 92-97, 98-2020

db <- "C:/Users/auerbdaa/Downloads/MSM_98-2020- SF2019_cmpct.mdb"
db_con <- DBI::dbConnect(
  drv = odbc::odbc(),
  .connection_string = paste0("Driver={Microsoft Access Driver (*.mdb, *.accdb)};DBQ=", db, ";")
  )

msm_tbl_names <- c("Edit_Distinct")

msm_tbl <- map(set_names(msm_tbl_names), ~collect(tbl(db_con, .x)))

msm_tbl$Edit_Distinct

DBI::dbDisconnect(db_con)

```

Starting to rough out an R version of MSM mapping...

```{r}
#Edit_Distinct.Gear == first character of RMIS fishery
#approx logic: check match of Edit_Distinct.PSC_Code against corresponding Num_Chars of recovery_location_code
#Edit_Distinct.Fishery_Number is a PSC fishery code, which then needs further mapping to FRAM

msm_tbl$Edit_Distinct |> count(Num_Chars)


split(msm_tbl$Edit_Distinct, msm_tbl$Edit_Distinct$Num_Chars)[1:2] |> 
  map(
    function(df){
      nc = df$Num_Chars[1]
      rmis_recs |> 
        mutate(
          PSC_Code = str_sub(recovery_location_code, 1, nc)
        ) |> 
        left_join(df, by = "PSC_Code")
    }
  )
  


```



# Recovery patterns

```{r}
#trial dummy example
mpdrec <- readxl::read_excel("T:/DFW-Salmon Mgmt Modeling Team - General/Southern Fund Projects/Coho FRAM MSM CWT analysis/MSMDB/MappedRecoveries.xlsx", sheet = "MappedRecoveries")

mpdrec |>
  ggplot(aes(RunYear)) +
  geom_bar()
#equivalently...
mpdrec |>
  count(RunYear) |> 
  ggplot(aes(RunYear, n)) +
  geom_col()


#all recoveries
mpdrec |>
  ggplot(aes(factor(RunYear), factor(PrNum))) +
  geom_bin2d()

#all non-escapement recoveries
mpdrec |>
  filter(MSMFishCode != 228) |> 
  ggplot(aes(factor(RunYear), factor(PrNum))) +
  geom_bin2d() +
  scale_fill_distiller(direction = 1)

```


## PR by year

## PR_MU by year


